Opening up the court (surface) in tennis grand slams

### 2019-09-05 paper_results/EDA/graphics_eda.Rmd
add graphs for the EDA section and for the complete data section

---
title: "Opening up the court (surface) in tennis grand slams"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
header-includes:
  - \usepackage{hyperref}
abstract: "IDK writing is hard"
---

@@

## To change into library(courtsports)
devtools::load_all("../..")

```{r graphing-parameters, include = FALSE}
# my_theme <-  theme_bw() + # White background, black and white theme
#   theme(axis.text = element_text(size = 12),
#         text = element_text(size = 14,
#         family="serif"),
#         plot.title = element_text(hjust = 0.5, size=16))

my_theme <-  theme_bw() + # White background, black and white theme
  theme(axis.text = element_text(size = 24),
        text = element_text(size = 28,
        family="serif"),
        plot.title = element_text(hjust = 0.5, size=34),
        plot.subtitle = element_text(hjust = 0.5))

@@

```{r modelling, eval = FALSE}
## HIERARCHICAL MODELLING
## NOT RUN
## WARNING: Takes about 1 HOUR

gs_players$name_int = as.integer(as.factor(gs_players$name))
gs_players$hand_fac = as.factor(gs_players$hand)
gs_players$ioc_fac = as.factor(gs_players$ioc)
gs_players$atp = gs_players$league == "ATP"
gs_players$year_fac = as.factor(gs_players$year)
gs_players$late_round = gs_players$round >= "R16"
gs_players$seeded = gs_players$rank <= 32
gs_players$opponent_seeded = gs_players$opponent_rank <= 32

set.seed(091418)
nocourt_logistic = glm(did_win ~ ioc_fac + late_round + log(rank) + log(opponent_rank) + year_fac + atp, data = gs_players, family = "binomial")
base_logistic = glm(did_win ~ ioc_fac +  tournament + late_round + log(rank) + log(opponent_rank) + year_fac + atp, data = gs_players, family = "binomial")
country_logistic = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |ioc_fac), data = gs_players, family = "binomial")
ind_logistic = glmer(did_win ~  ioc_fac + late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name_int), data = gs_players, family = "binomial")
ind_logistic_noioc = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name_int), data = gs_players, family = "binomial")
ind_int_logistic = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + tournament + (1|name_int), data = gs_players, family = "binomial")
ind_year_logistic = glmer(did_win ~ late_round + log(rank) + log(opponent_rank) + atp + tournament + (0+year_fac|name_int), data = gs_players, family = "binomial")
n_aces_mod = lmer(n_aces ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name), data = gs_partial_players)
n_winners_mod = lmer(n_winners ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name), data = gs_partial_players)
n_net_mod = lmer(n_netpt_w ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name), data = gs_partial_players)
n_ue_mod = lmer(n_ue ~ late_round + log(rank) + log(opponent_rank) + year_fac + atp + (0 + tournament |name), data = gs_partial_players)

# Data and EDA {#sec:data-eda}

## Data {#sec:data}

```{r tab-data, include = TRUE, echo = FALSE}
w_id <- which(gs$winner_name == "Serena Williams")[1:2]
f_id <- which(gs$winner_name == "Roger Federer")[1:2]
n_id <- which(gs$winner_name == "Rafael Nadal")[1:2]
s_id <- which(gs$winner_name %in% c("Serena Williams", "Roger Federer", "Rafael Nadal"))
cols <- c("winner_name", "tourney_name", "year", "winner_ioc", "w_pointswon", "winner_rank", "l_pointswon", "loser_rank")
knitr::kable(gs[c(w_id, f_id, n_id), cols], format = "latex",
             row.names = FALSE, col.names = c("Winner", "Tournament", "Year", "W. IOC", "W. Points",  "W. Rank", "L. Points", "L. Rank"),
             caption = "\\label{tab:data}Example of the complete grand slam data.  It includes winner and loser attributes, match attributes, and tournament attributes.  Not all attributes are shown here.",
             booktabs = TRUE) %>% kable_styling(latex_options = "striped", "hold_position")
```

## Early Data Analysis {#sec:eda}

### Examining the distribution of points earned

```{r, fig.height=5, fig.width=10}

# grid.arrange(p1, p2, p3, ncol=3, widths=c(.25, .41, .34))

p2 <- ggplot(gs_players) + 
  geom_density(aes(pointswon, ..scaled..,
                   color=tournament, group=tournament, fill = tournament),
                                        size = 3, alpha = 0) + 
  scale_fill_manual(" ", values=tournament_colors,
                                labels = c("Australian Open (Plexicushion)",
                                           "French Open (Clay)",
                                           "US Open (DecoTurf)",
                                           "Wimbledon (Grass)")) + my_theme  +
  guides(fill = guide_legend(override.aes = list(size = 0, alpha = 1),
                              nrow=4, byrow=TRUE)) +
    theme(legend.position="bottom") +
    scale_color_manual(guide=FALSE, values=tournament_colors) +
  facet_wrap(~league, ncol=1) +
  labs(x = "Points", y = "Scaled density")

# 
# ggplot(gs_players) + geom_density(aes(pointswon, color=tournament, 
#                                             group=tournament, fill = tournament),
#                                         size = 3, alpha = 0) + 
#   scale_fill_manual(" ", values=tournament_colors,
#                                 labels = c("Australian Open\n(Plexicushion)",
#                                            "French Open\n(Clay)",
#                                            "US Open\n(DecoTurf)",
#                                            "Wimbledon\n(Grass)")) + my_theme  +
#   guides(fill = guide_legend(override.aes = list(size = 0, alpha = 1),
#                               nrow=2, byrow=TRUE)) +
#     theme(legend.position="bottom") +
#     scale_color_manual(guide=FALSE, values=tournament_colors)

@@

grid.arrange(p2, p1, ncol=2) %>%
             # # top = textGrob("Distribution of points per match",
             # #                gp=gpar(fontsize=20, fontfamily="serif")),
             #  layout_matrix = rbind(c(2, 1),
             #                        c(2, 1),
             #                        c(NA, 1)),
             # heights=c(1, 3, .1)) %>%
  ggsave("plots/points_per_match.jpg", ., height=8, width=graphic_width_long)

# 
# ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom") %>%
# grid.arrange(top = textGrob("Spaniards perform better on Clay: confirmed?",
#                             gp=gpar(fontsize=20, fontfamily="serif")),
#              bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
#                            # x = 0,
#                            # y = 0.5,
#                            just = "left",
#                            gp=gpar(fontfamily="serif")))
# 
# ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom") %>%
#   ggsave("plots/spainards.jpg", ., height=7, width=16)

### Home court advantage

```{r, fig.height=3.5, fig.width=10, fig.align='center'}
france_wins <- dplyr::filter(gs_players, ioc == "FRA", round < "QF") %>% 
  dplyr::group_by(tournament, round) %>% 
  dplyr::summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

@@

our_alpha <- 0.5

p1<- ggplot(france_wins, aes(round, win_prop,
                             color=tournament, group=tournament,
                             alpha = tournament,
                             size = tournament)) + 
  # geom_line(size = 3) + 
  # geom_point() +
  labs(title = "French players",
       x = "", y = " ") + 
  geom_line(size=3) + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
scale_alpha_manual(guide = FALSE,
                   values = c(our_alpha, 1, our_alpha, our_alpha)) + 
scale_size_manual(guide = FALSE,
                   values = c(1, 3, 1, 1))

p2 <- ggplot(us_wins, aes(round, win_prop,
                          color=tournament, group=tournament,
                          alpha = tournament)) + 
  # geom_line() + 
  # geom_point() +
  labs(title = "USA players",
              x = "", y = " ") + 
       # x = "", y = "Win proportion") + 
    geom_line(size=3) + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
scale_alpha_manual(guide = FALSE,
                   values = c(our_alpha, our_alpha, 1, our_alpha))

p3 <- ggplot(aus_wins, aes(round, win_prop,
                           color=tournament, group=tournament, alpha = tournament)) + 
  # geom_line() + 
  # geom_point() +
  labs(title = "Aus. players",
       x = "", y = "Win proportion") + 
         # x = "", y = " ") + 
    geom_line(size=3) + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
scale_alpha_manual(guide = FALSE,
                   values = c(1, our_alpha, our_alpha, our_alpha))

p4 <- ggplot(gbr_wins, aes(round, win_prop, color=tournament,
                           group=tournament, fill = tournament,
                           alpha = tournament)) + 
  geom_line(size=3) + 
  labs(title = "British players",
       # x = "", y = "Win proportion") + 
       x = "", y = " ") + 
  scale_color_manual(" ", values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme +
  guides(color = guide_legend(override.aes = list(size = 7)))  + 
scale_alpha_manual(guide = FALSE,
                   values = c(our_alpha, our_alpha, our_alpha, 1))

# 
# + 
#   geom_line() + 
#   # geom_point() +
#   labs(title = "British players",
#        x = "", y = "Win proportion") + 
#   scale_color_manual(" ", values=tournament_colors) + my_theme

ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(., top = textGrob("Home court advantage: debunked",
                            gp=gpar(fontsize=20, fontfamily="serif")),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif")))

ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             # top = textGrob("Home court advantage: debunked",
             #                gp=gpar(fontsize=20, fontfamily="serif")),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("plots/home_court.jpg", ., height=5, width=graphic_width_long)

@@

```{r}

dplyr::filter(gs_players, ioc == "USA", round < "QF")  %>% nrow()
dplyr::filter(gs_players, ioc == "USA", round < "QF")  %>% pull(name) %>% unique() %>% length()

@@

dplyr::filter(gs_players, ioc == "GBR", round < "QF")  %>% pull(name) %>% unique() %>% length()
dplyr::filter(gs_players, ioc == "GBR", round < "QF")  %>% nrow()

p1<- ggplot(france_wins, aes(round, win_prop,
                             color=tournament, group=tournament,
                             # alpha = tournament,
                             size = tournament,
                             linetype = tournament)) + 
  # geom_line(size = 3) + 
  # geom_point() +
  labs(title = "French players",
       subtitle = "53 players, n=828",
       x = "", y = " ") + 
  geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
# scale_alpha_manual(guide = FALSE,
#                    values = c(our_alpha, 1, our_alpha, our_alpha)) + 
scale_size_manual(guide = FALSE,
                   values = c(1, 3, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 1, 2, 2)) + ylim(c(0, 1))

p2 <- ggplot(us_wins, aes(round, win_prop,
                          color=tournament, group=tournament,
                          # alpha = tournament,
                          size = tournament,
                          linetype=tournament)) + 
  # geom_line() + 
  # geom_point() +
  labs(title = "USA players",
       subtitle = "81 players, n=1046",
              x = "", y = " ") + 
       # x = "", y = "Win proportion") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 3, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 1, 2)) + ylim(c(0, 1))
# scale_alpha_manual(guide = FALSE,
#                    values = c(our_alpha, our_alpha, 1, our_alpha))

p3 <- ggplot(aus_wins, aes(round, win_prop,
                           color=tournament, group=tournament, # alpha = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  # geom_line() + 
  # geom_point() +
  labs(title = "Aus. players",
       subtitle = "40 players, n=376",
       x = "", y = "Win proportion") + 
         # x = "", y = " ") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(3, 1, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(1, 2, 2, 2)) + ylim(c(0, 1))
# scale_alpha_manual(guide = FALSE,
#                    values = c(1, our_alpha, our_alpha, our_alpha))

p4 <- ggplot(gbr_wins, aes(round, win_prop, color=tournament,
                           group=tournament, fill = tournament,
                           # alpha = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  geom_line() + 
  labs(title = "British players",
       subtitle = "21 players, n=249",
       # x = "", y = "Win proportion") + 
       x = "", y = " ") + 
  scale_color_manual(" ", values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme +
  guides(color = guide_legend(override.aes = list(size = 7)))  + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 1, 3)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 2, 1)) + ylim(c(0, 1))

# scale_alpha_manual(guide = FALSE,
#                    values = c(our_alpha, our_alpha, our_alpha, 1))

ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             # top = textGrob("Home court advantage: debunked",
             #                gp=gpar(fontsize=20, fontfamily="serif")),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("plots/home_court.jpg", ., height=5.5, width=graphic_width_long)

```{r}
# home court only with those in all 4

all4 <- 
  gs_players %>% 
  group_by(name) %>% 
  summarize(blah = length(unique(tournament))) %>%
  filter(blah == 4) %>%
  pull(name)

@@

p1<- ggplot(france_wins, aes(round, win_prop,
                             color=tournament, group=tournament,
                             # alpha = tournament,
                             size = tournament,
                             linetype = tournament)) + 
  # geom_line(size = 3) + 
  # geom_point() +
  labs(title = "French players",
       x = "", y = " ") + 
  geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
# scale_alpha_manual(guide = FALSE,
#                    values = c(our_alpha, 1, our_alpha, our_alpha)) + 
scale_size_manual(guide = FALSE,
                   values = c(1, 3, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 1, 2, 2)) + ylim(c(0, 1))

p2 <- ggplot(us_wins, aes(round, win_prop,
                          color=tournament, group=tournament,
                          # alpha = tournament,
                          size = tournament,
                          linetype=tournament)) + 
  # geom_line() + 
  # geom_point() +
  labs(title = "USA players",
              x = "", y = " ") + 
       # x = "", y = "Win proportion") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 3, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 1, 2)) + ylim(c(0, 1))
# scale_alpha_manual(guide = FALSE,
#                    values = c(our_alpha, our_alpha, 1, our_alpha))

p3 <- ggplot(aus_wins, aes(round, win_prop,
                           color=tournament, group=tournament, # alpha = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  # geom_line() + 
  # geom_point() +
  labs(title = "Aus. players",
       x = "", y = "Win proportion") + 
         # x = "", y = " ") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(3, 1, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(1, 2, 2, 2)) + ylim(c(0, 1))
# scale_alpha_manual(guide = FALSE,
#                    values = c(1, our_alpha, our_alpha, our_alpha))

p4 <- ggplot(gbr_wins, aes(round, win_prop, color=tournament,
                           group=tournament, fill = tournament,
                           # alpha = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  geom_line() + 
  labs(title = "British players",
       # x = "", y = "Win proportion") + 
       x = "", y = " ") + 
  scale_color_manual(" ", values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme +
  guides(color = guide_legend(override.aes = list(size = 7)))  + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 1, 3)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 2, 1)) + ylim(c(0, 1))

# scale_alpha_manual(guide = FALSE,
#                    values = c(our_alpha, our_alpha, our_alpha, 1))

@@

ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             # top = textGrob("Only those who made all 4",
             #                gp=gpar(fontsize=20, fontfamily="serif")),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("plots/home_court_all4.jpg", ., height=5, width=graphic_width_long)

@@

```{r}

spain_wins <- filter(gs_players, ioc == "ESP", round < "QF") %>% 
  group_by(tournament, round) %>% 
  dplyr::summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()),
            avg_ranking = mean(rank),
            rank_se = sd(rank)/sqrt(n()))

@@

################################################################################
# TESTS FOR PAPER
################################################################################
with(filter(gs_players, ioc == "ESP", round == "R128"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
with(filter(gs_players, ioc == "ESP", round == "R64"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
with(filter(gs_players, ioc == "ESP", round == "R32"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
with(filter(gs_players, ioc == "ESP", round == "R16"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
```

```{r, fig.height=3.8, fig.width=8, out.width="80%", fig.align='center', fig.cap = "\\label{fig:spain-clay}"}

# my_theme <-  theme_bw() + # White background, black and white theme
#   theme(axis.text = element_text(size = 24),
#         text = element_text(size = 28,
#         family="serif"),
#         plot.title = element_text(hjust = 0.5, size=34))

p1 <- ggplot(spain_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  labs(title = "Win proportion among Spaniards",
       # caption = "Note: we only look at the first 4 rounds due to decreasing sample size",
       x = "", y = "Win proportion") + 
    geom_line(size=3) + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme

@@

ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom") %>%
grid.arrange(top = textGrob("Spaniards perform better on Clay: confirmed?",
                            gp=gpar(fontsize=20, fontfamily="serif")),
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           # x = 0,
                           # y = 0.5,
                           just = "left",
                           gp=gpar(fontfamily="serif")))

ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom") %>%
  ggsave("plots/spainards.jpg", ., height=6, width=graphic_width_long)

@@

# Unforced errors and aces

>>>>

---
title: "Opening up the court (surface) in tennis grand slams"
date: ' '
output:
  pdf_document:
    toc:  true
header-includes:
  - \usepackage{hyperref}
---

@@

## To change into library(courtsports)
devtools::load_all("../..")
```

```{r graphing-parameters, include = FALSE}
my_theme <-  theme_bw() + # White background, black and white theme
  theme(axis.text = element_text(size = 24),
        text = element_text(size = 28,
        family="serif"),
        plot.title = element_text(hjust = 0.5, size=34),
        plot.subtitle = element_text(hjust = 0.5))

@@

# SHOULD WE WRITE/SAVE THE GRAPHICS LOCALLY??
save_graph <- FALSE
```

# EDA

## Distribution of points earned

```{r, fig.height=8, fig.width=16}
p2 <- ggplot(gs_players) + 
  geom_density(aes(pointswon, ..scaled..,
                   color=tournament, group=tournament, fill = tournament),
                                        size = 3, alpha = 0) + 
  scale_fill_manual(" ", values=tournament_colors,
                                labels = c("Australian Open (Plexicushion)",
                                           "French Open (Clay)",
                                           "US Open (DecoTurf)",
                                           "Wimbledon (Grass)")) + my_theme  +
  guides(fill = guide_legend(override.aes = list(size = 0, alpha = 1),
                              nrow=4, byrow=TRUE)) +
    theme(legend.position="bottom") +
    scale_color_manual(guide=FALSE, values=tournament_colors) +
  facet_wrap(~league, ncol=1) +
  labs(x = "Points", y = "Scaled density")

@@

grid.arrange(p2, p1, ncol=2)

if(save_graph){
  grid.arrange(p2, p1, ncol=2) %>%
  ggsave("plots/points_per_match.jpg", ., height=8, width=graphic_width_long)
}
```

## Home court advantage

```{r}
france_wins <- dplyr::filter(gs_players, ioc == "FRA", round < "QF") %>% 
  dplyr::group_by(tournament, round) %>% 
  dplyr::summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()))

@@

@@

```{r, echo=TRUE}
dplyr::filter(gs_players, ioc == "USA", round < "QF")  %>% nrow()
dplyr::filter(gs_players, ioc == "USA", round < "QF")  %>% 
  pull(name) %>% unique() %>% length()

@@

dplyr::filter(gs_players, ioc == "GBR", round < "QF")  %>% nrow()
dplyr::filter(gs_players, ioc == "GBR", round < "QF")  %>% 
  pull(name) %>% unique() %>% length()
```

```{r, fig.height=5.5, fig.width=16, fig.align='center'}
p1<- ggplot(france_wins, aes(round, win_prop,
                             color=tournament, group=tournament,
                             size = tournament,
                             linetype = tournament)) + 
  labs(title = "French players",
       subtitle = "53 players, n=828",
       x = "", y = " ") + 
  geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
scale_size_manual(guide = FALSE,
                   values = c(1, 3, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 1, 2, 2)) + ylim(c(0, 1))

p2 <- ggplot(us_wins, aes(round, win_prop,
                          color=tournament, group=tournament,
                          size = tournament,
                          linetype=tournament)) + 
  labs(title = "USA players",
       subtitle = "81 players, n=1046",
              x = "", y = " ") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 3, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 1, 2)) + ylim(c(0, 1))

p3 <- ggplot(aus_wins, aes(round, win_prop,
                           color=tournament, group=tournament, # alpha = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  labs(title = "Aus. players",
       subtitle = "40 players, n=376",
       x = "", y = "Win proportion") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(3, 1, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(1, 2, 2, 2)) + ylim(c(0, 1))

p4 <- ggplot(gbr_wins, aes(round, win_prop, color=tournament,
                           group=tournament, fill = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  geom_line() + 
  labs(title = "British players",
       subtitle = "21 players, n=249",
       x = "", y = " ") + 
  scale_color_manual(" ", values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme +
  guides(color = guide_legend(override.aes = list(size = 7)))  + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 1, 3)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 2, 1)) + ylim(c(0, 1))

ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) 
if(save_graph){
ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("plots/home_court.jpg", ., height=5.5, width=graphic_width_long)
}
```

### Subset: players who played in all 4 tournaments

```{r, fig.height=5, fig.width=16, fig.align='center'}
all4 <- 
  gs_players %>% 
  group_by(name) %>% 
  summarize(blah = length(unique(tournament))) %>%
  filter(blah == 4) %>%
  pull(name)

@@

p1<- ggplot(france_wins, aes(round, win_prop,
                             color=tournament, group=tournament,
                             size = tournament,
                             linetype = tournament)) + 
  labs(title = "French players",
       x = "", y = " ") + 
  geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
scale_size_manual(guide = FALSE,
                   values = c(1, 3, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 1, 2, 2)) + ylim(c(0, 1))

p2 <- ggplot(us_wins, aes(round, win_prop,
                          color=tournament, group=tournament,
                          size = tournament,
                          linetype=tournament)) + 
  labs(title = "USA players",
              x = "", y = " ") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 3, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 1, 2)) + ylim(c(0, 1))

p3 <- ggplot(aus_wins, aes(round, win_prop,
                           color=tournament, group=tournament, # alpha = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  labs(title = "Aus. players",
       x = "", y = "Win proportion") + 
    geom_line() + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme   + 
  scale_size_manual(guide = FALSE,
                   values = c(3, 1, 1, 1)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(1, 2, 2, 2)) + ylim(c(0, 1))

p4 <- ggplot(gbr_wins, aes(round, win_prop, color=tournament,
                           group=tournament, fill = tournament,
                           size = tournament,
                           linetype = tournament)) + 
  geom_line() + 
  labs(title = "British players",
       x = "", y = " ") + 
  scale_color_manual(" ", values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme +
  guides(color = guide_legend(override.aes = list(size = 7)))  + 
  scale_size_manual(guide = FALSE,
                   values = c(1, 1, 1, 3)) + 
  scale_linetype_manual(guide = FALSE,
                   values = c(2, 2, 2, 1)) + ylim(c(0, 1))

@@

if(save_graph){
ggarrange(p3, p1, p2,  p4, nrow=1, ncol=4, common.legend = TRUE, legend="bottom") %>%
grid.arrange(.,
             bottom = textGrob("Note: we only look at the first 4 rounds due to decreasing sample size",
                           just = "left",
                           gp=gpar(fontfamily="serif", fontsize = 20))) %>%
  ggsave("plots/home_court_all4.jpg", ., height=5, width=graphic_width_long)
}
```

@@

```{r}
spain_wins <- filter(gs_players, ioc == "ESP", round < "QF") %>% 
  group_by(tournament, round) %>% 
  dplyr::summarise(win_prop = mean(did_win),
            win_prop_se = sqrt(win_prop*(1-win_prop)/n()),
            avg_ranking = mean(rank),
            rank_se = sd(rank)/sqrt(n()))

@@

### Chi-squared tests

```{r}
################################################################################
# TESTS FOR PAPER
################################################################################
with(filter(gs_players, ioc == "ESP", round == "R128"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
with(filter(gs_players, ioc == "ESP", round == "R64"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
with(filter(gs_players, ioc == "ESP", round == "R32"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
with(filter(gs_players, ioc == "ESP", round == "R16"),
     table(did_win, tournament)) %>%
  chisq.test(correct = TRUE)
```

```{r, fig.height=6, fig.width=16, fig.align='center'}
p1 <- ggplot(spain_wins, aes(round, win_prop, color=tournament, group=tournament)) + 
  labs(title = "Win proportion among Spaniards",
       x = "", y = "Win proportion") + 
    geom_line(size=3) + 
  scale_color_manual(guide=FALSE, values=tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)",
                                           "French Open\n(Clay)",
                                           "US Open\n(DecoTurf)",
                                           "Wimbledon\n(Grass)")) + my_theme

@@

ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom")

if(save_graph){
ggarrange( p1, p2, nrow=1, ncol=2, common.legend = TRUE, legend="bottom") %>%
  ggsave("plots/spainards.jpg", ., height=6, width=graphic_width_long)
}

@@

## Unforced errors and aces
